---
title: "Lab 1: Intro to R"
author: "RickRN"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

Following is the vector of girls born during the years 1629-1710.

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

There is an increasing trend in the number of girls born in the years of 1629-1710.

```{r trend-girls}
# Insert code for Exercise 2 here
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()

```


### Exercise 3

There are slightly more boys than girls born in the years of 1629-1710.

```{r plot-prop-boys-arbuthnot}
# Insert code for Exercise 3 here
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line()

```


### Exercise 4

Following are the dimensions of the "present" birth records in the United States.

```{r dim-present}
# Insert code for Exercise 4 here
dim(present)
```



### Exercise 5

The 'present' dataset has 19 less cases ("years") than the arbuthnot dataset.

```{r count-compare}
# Insert code for Exercise 5 here
present %>% summarize(min = min(year), max(year))

arbuthnot %>% summarize(min = min(year), max(year))

cat("Present: ", + length(present$year))
cat(" Arbuthnot: ", + length(arbuthnot$year))

cat(" Present difference in cases: ", + (case_diff <- length(present$year) - length(arbuthnot$year)))
    


```


### Exercise 6

There are slightly more boys than girls born in the years of 1940-2002. However, this is a negative trend.


```{r plot-prop-boys-present}
# Insert code for Exercise 6 here
present <- present %>%
  mutate(total = boys + girls)

present <- present %>%
  mutate(boy_ratio = boys / total)

ggplot(data = present, aes(x = year, y = boy_ratio)) + 
  geom_line()

```


### Exercise 7

The year 1961 had the most total number of births in the US at 4,268,326.

```{r find-max-total}
# Insert code for Exercise 7 here
present %>%
  arrange(desc(total))

```

